Revealing transient strain in geodetic data with Gaussian process regression

### 2017-10-11 RevisedSubmission/hines_hetland_2017_gji.tex
did more

Most methods for estimating strain rates from GNSS data assume some parametric form of the deformation signal. The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first-order polynomial (i.e., having constant deformation gradients) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}. The components of the strain rate tensor  for each subnetwork are then determined through a least squares fit to the observations. The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span too large of an area. To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at some point is primarily controlled by observations at nearby stations. However, the methods described in \citet{Shen1996,Shen2015} are still formulated by assuming that the deformation gradients are uniform over the entire network. The errors in this assumption manifest as implausibly low formal uncertainties for the estimated strain rates. Other methods for estimating secular strain rates have parameterized GNSS derived velocities with bi-cubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}. The type of basis functions and the number of degrees of freedom for a parameterization can be subjective. If there are too few degrees of freedom in the parameterization, then estimated strain rates will be biased and the uncertainties will be underestimated. On the other hand, if there are too many degrees of freedom, then there will not be any coherent features in the estimated strain rates. The methods described by \citet{Beavan2001} and \citet{Tape2009} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution. One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}. In such models the lithospheric rheology and fault geometries are assumed to be known. Any errors in the assumed physical model could result in biased strain estimates and underestimated formal uncertainties.

>>>>

Most methods for estimating strain rates from GNSS data assume some parametric form for the deformation signal. The simplest method for estimating secular strain rates assumes that GNSS derived velocities can be described with a first-order polynomial (i.e, the deformation gradients are constant) over some subnetwork of the GNSS stations \citep[e.g.,][]{Feigl1990,Murray2000}. The components of the strain rate tensor for each subnetwork are then determined from the least squares fit to the observations. The assumption that deformation gradients are spatially uniform is not appropriate when subnetworks span too large of an area. To help overcome this deficiency, \citet{Shen1996,Shen2015} used an inverse distance weighting scheme, in which the estimated strain rate at a point is primarily controlled by observations at nearby stations. The method  of \citet{Shen1996,Shen2015} can be viewed as a form of local least squares regression with a first-order polynomial \citep[e.g.,][sec. 6]{Hastie2009}. Other methods for estimating secular strain rates have parameterized GNSS derived velocities with bi-cubic splines \citep{Beavan2001}, spherical wavelets \citep{Tape2009}, and elastostatic Green's functions \citep{Sandwell2016}. The type of basis functions and the number of degrees of freedom for a parameterization, which are often chosen subjectively, have a strong influence on the strain solution. If there are too few degrees of freedom in the parameterization, then estimated strain rates will be biased and the uncertainties will be underestimated. On the other hand, if there are too many degrees of freedom, then there will not be any coherent features in the estimated strain rates. The methods described by \citet{Beavan2001} and \citet{Tape2009} also require the user to specify penalty parameters that control a similar trade-off between bias and variance in the solution. One could parameterize deformation with a physically motivated model of interseismic deformation \citep[e.g.,][]{Meade2005,McCaffrey2007}. In such models the lithospheric rheology and fault geometries are assumed to be known. Any errors in the assumed physical model could result in biased strain estimates and underestimated formal uncertainties.
